---
title: "Phase 1: Simulation and OLS parameter recovery"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
---

## Purpose

Before fitting any hierarchical Bayesian model to market data, verify that the data-generating process is internally consistent and that a conventional OLS estimator recovers the known stock-level factor betas as information increases.

This notebook is intentionally simple. It is the baseline against which ridge, independent Bayesian, and hierarchical Bayesian estimators will be compared.

```{r}
source("../R/simulation.R")
source("../R/model_ols.R")
source("../R/evaluation.R")
```

## One simulated panel

```{r}
sim <- simulate_factor_panel(
  n_stocks = 100,
  n_sectors = 10,
  n_obs = 156,
  seed = 123
)

str(sim$metadata)
head(sim$truth$stock)
```

The hierarchy is known by construction:

```text
market factor mean
      |
sector factor beta
      |
 stock factor beta
      |
 observed return
```

## OLS recovery

```{r}
ols <- fit_ols_betas(sim$panel, sim$metadata$factor_names)
recovery <- join_estimates_to_truth(ols, sim$truth$stock)
summarise_beta_recovery(recovery)
```

## Estimated versus true betas

```{r}
plot(
  recovery$beta,
  recovery$estimate,
  xlab = "True beta",
  ylab = "OLS estimate",
  main = "OLS recovery across stocks and factors"
)
abline(a = 0, b = 1)
```

## Data-scarcity baseline

The key future experiment asks whether hierarchical shrinkage becomes more valuable as the stock-level sample becomes shorter. First establish the OLS error curve.

```{r}
ols_curve <- run_ols_recovery_experiment(
  sample_sizes = c(26, 52, 104, 156, 260, 520),
  n_stocks = 100,
  n_sectors = 10,
  seed = 123
)

ols_all <- ols_curve[ols_curve$factor == "ALL", ]
ols_all
```

```{r}
plot(
  ols_all$n_obs,
  ols_all$rmse,
  type = "b",
  log = "x",
  xlab = "Weekly observations (log scale)",
  ylab = "Beta RMSE",
  main = "OLS beta error versus available history"
)
```

## Acceptance criteria for this phase

- simulation is deterministic under a fixed seed;
- dimensions and stock-sector mapping are correct;
- every estimated beta joins to exactly one true beta;
- OLS beta RMSE generally falls as the available history increases;
- this notebook can be reproduced before any Bayesian model is introduced.

The next estimator to add is ridge regression. Hierarchical Bayes will only be judged against OLS **and** a non-Bayesian shrinkage benchmark.
