---
title: "OLS vs Ridge: Parameter-Recovery Baselines"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Research question

Before introducing hierarchical Bayes, test whether generic shrinkage already improves stock-level factor-beta estimation.

The comparison is deliberately narrow:

1. stock-by-stock OLS;
2. stock-by-stock ridge regression;
3. identical simulated panels with known true betas;
4. a fixed sample-size grid from 26 to 520 weekly observations.

Ridge tuning is performed with expanding-window validation. Predictor centering and scaling are estimated inside each training fold only.

```{r}
source("R/simulation.R")
source("R/model_ols.R")
source("R/model_ridge.R")
source("R/evaluation.R")

sample_sizes <- c(26L, 52L, 104L, 156L, 260L, 520L)
lambda_grid <- ridge_lambda_grid(n = 21L)
```

## One-panel sanity check

```{r}
sim <- simulate_factor_panel(
  n_stocks = 100L,
  n_sectors = 10L,
  n_obs = 156L,
  seed = 123L
)

ols <- fit_ols_betas(sim$panel, sim$metadata$factor_names)
ridge <- fit_ridge_betas(
  sim$panel,
  sim$metadata$factor_names,
  lambda_grid = lambda_grid
)

ols_joined <- join_estimates_to_truth(ols, sim$truth$stock)
ridge_joined <- join_estimates_to_truth(ridge, sim$truth$stock)

ols_summary <- summarise_beta_recovery(ols_joined)
ridge_summary <- summarise_beta_recovery(ridge_joined)

ols_summary$estimator <- "OLS"
ridge_summary$estimator <- "Ridge"
rbind(ols_summary, ridge_summary)
```

## How much shrinkage is being selected?

```{r}
stock_lambdas <- unique(ridge[, c("stock", "lambda", "cv_mse")])
summary(stock_lambdas$lambda)
```

The selected lambda is allowed to differ across stocks because the signal-to-noise ratio differs across return series. The factor design matrix is shared in the simulation, but the response noise is stock-specific.

## Data-scarcity experiment

```{r}
baselines <- run_baseline_recovery_experiment(
  sample_sizes = sample_sizes,
  n_stocks = 100L,
  n_sectors = 10L,
  seed = 123L,
  lambda_grid = lambda_grid
)

baselines
```

## Overall recovery curve

```{r}
overall <- baselines[baselines$factor == "ALL", ]

plot(
  overall$n_obs[overall$estimator == "OLS"],
  overall$rmse[overall$estimator == "OLS"],
  type = "b",
  xlab = "Weekly observations available",
  ylab = "Beta RMSE",
  main = "Parameter recovery: OLS vs ridge"
)
lines(
  overall$n_obs[overall$estimator == "Ridge"],
  overall$rmse[overall$estimator == "Ridge"],
  type = "b",
  lty = 2
)
legend("topright", legend = c("OLS", "Ridge"), lty = c(1, 2), pch = 1)
```

## Factor-by-factor recovery

```{r}
factor_rows <- baselines[baselines$factor != "ALL", ]
wide <- reshape(
  factor_rows[, c("estimator", "n_obs", "factor", "rmse")],
  idvar = c("n_obs", "factor"),
  timevar = "estimator",
  direction = "wide"
)
wide$ridge_minus_ols_rmse <- wide$rmse.Ridge - wide$rmse.OLS
wide[order(wide$n_obs, wide$factor), ]
```

Negative values of `ridge_minus_ols_rmse` indicate that ridge recovers the known beta more accurately than OLS.

## Interpretation rule

This notebook is a benchmark, not a demonstration that ridge must win. If ridge fails to improve recovery, that is a result. The later hierarchical Bayesian model must beat both OLS and ridge on the same simulated designs before claims about economically structured pooling are credible.
