---
title: "Hierarchical Bayesian Factor Model: Prior Predictive Protocol"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Objective

Before looking at posterior parameter recovery, verify that the prior implies plausible weekly stock returns and plausible cross-sectional variation in factor exposures.

The v0.1 hierarchy is

\[
\beta_{i,k} = \beta_{0,k} + u_{s(i),k} + v_{i,k},
\]

with

\[
u_{s,k} \sim N(0, \tau_{sector,k}^2),
\qquad
v_{i,k} \sim N(0, \tau_{stock,k}^2).
\]

The first implementation uses a common sector-level standard deviation and a common stock-level standard deviation for each factor across all sectors. It does **not** estimate a separate `tau` for every sector. That more flexible model should only be introduced if posterior predictive checks or out-of-sample evidence show that the simpler hierarchy is inadequate.

Group-level correlations are also disabled initially. This makes the first research question clean: does partial pooling of individual factor slopes help?

```{r}
source("R/simulation.R")
source("R/model_bayes.R")

sim <- simulate_factor_panel(
  n_stocks = 40L,
  n_sectors = 8L,
  n_obs = 104L,
  seed = 123L
)

design <- validate_hierarchical_design(
  sim$panel,
  sim$metadata$factor_names
)
design

make_hierarchical_factor_formula(sim$metadata$factor_names)
```

## Initial prior family

For the first fit, the intended scales are:

- population factor slopes: `Normal(0, 1)`;
- population intercept: `Normal(0, 0.01)`;
- sector and stock factor-slope standard deviations: half-`Normal(0, 0.35)`;
- sector and stock intercept standard deviations: half-`Normal(0, 0.01)`;
- residual volatility: half-`Student-t(3, 0, 0.03)`.

These are starting points, not sacred constants. They must survive prior-predictive simulation.

## Prior predictive fit

The following is intentionally separated from posterior fitting.

```{r eval=false}
prior_fit <- fit_hierarchical_prior_predictive(
  panel = sim$panel,
  factor_names = sim$metadata$factor_names,
  chains = 4L,
  iter = 1000L,
  seed = 123L,
  backend = "cmdstanr"
)

brms::pp_check(prior_fit, ndraws = 100)
```

## Prior-predictive rejection rules

Do **not** proceed to posterior recovery merely because Stan samples successfully. Revise the prior if it puts substantial mass on obviously implausible weekly outcomes.

For the synthetic design, inspect at least:

1. the distribution of simulated weekly returns;
2. the implied cross-sectional spread of market betas;
3. the implied cross-sectional spread of the smaller style-factor betas;
4. the implied residual volatility;
5. whether prior draws routinely imply absurd multi-standard-deviation weekly returns.

The exact cutoffs should be documented after seeing the scale of the simulated inputs, but they must be chosen for economic plausibility rather than to improve posterior results.

## First posterior recovery fit

Only after the prior-predictive check passes:

```{r eval=false}
fit <- fit_hierarchical_bayes(
  panel = sim$panel,
  factor_names = sim$metadata$factor_names,
  chains = 4L,
  iter = 2000L,
  warmup = 1000L,
  seed = 123L,
  backend = "cmdstanr",
  adapt_delta = 0.95,
  max_treedepth = 12L
)

summary(fit)
brms::pp_check(fit)
```

## Diagnostics required before interpreting shrinkage

A fit is not accepted solely because posterior means look reasonable. Record:

- split R-hat, with a target below 1.01;
- bulk ESS;
- tail ESS;
- divergent transitions;
- maximum-treedepth warnings;
- posterior predictive checks;
- parameter recovery against the known stock betas;
- interval coverage;
- sensitivity to at least one narrower and one wider prior scale.

## Core comparison

The posterior recovery experiment must eventually report the same schema for:

1. OLS;
2. ridge;
3. independent Bayesian regression;
4. hierarchical Bayesian regression.

The hierarchical model earns its extra complexity only if it improves recovery or uncertainty calibration beyond generic ridge shrinkage.
