---
title: "Parameter Recovery: OLS, Ridge, Independent Bayes, Hierarchical Bayes"
format:
  html:
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Purpose

This notebook is the first experiment that can justify the hierarchical model.

The estimators are:

1. OLS;
2. ridge with expanding-window tuning;
3. independent Bayesian regression with no information sharing across stocks;
4. hierarchical Bayesian regression with market -> sector -> stock partial pooling.

All are scored against known simulated stock betas.

```{r}
source("R/simulation.R")
source("R/model_ols.R")
source("R/model_ridge.R")
source("R/model_bayes.R")
source("R/hierarchy_stress.R")
source("R/evaluation.R")

factor_names <- c("MKT", "SMB", "HML", "RMW", "CMA", "MOM")
```

## Start small

Do not begin with 100 stocks and the full sample-size grid. First prove that the entire pipeline works on a small synthetic panel.

```{r}
sim <- simulate_hierarchy_scenario(
  scenario = "balanced",
  n_stocks = 20L,
  n_sectors = 4L,
  n_obs = 104L,
  seed = 123L
)
```

## Frequentist baselines

```{r}
ols <- fit_ols_betas(sim$panel, factor_names)
ridge <- fit_ridge_betas(
  sim$panel,
  factor_names,
  lambda_grid = ridge_lambda_grid(n = 21L)
)

ols_joined <- join_estimates_to_truth(ols, sim$truth$stock)
ridge_joined <- join_estimates_to_truth(ridge, sim$truth$stock)

ols_recovery <- summarise_beta_recovery(ols_joined)
ridge_recovery <- summarise_beta_recovery(ridge_joined)
```

## Prior predictive check first

Do not execute posterior fits until the prior-predictive notebook has been reviewed.

```{r eval=false}
prior_fit <- fit_hierarchical_prior_predictive(
  sim$panel,
  factor_names,
  chains = 4L,
  iter = 1000L,
  seed = 123L
)

brms::pp_check(prior_fit, ndraws = 100)
```

## Independent Bayesian benchmark

For the first end-to-end test, fit only 20 stocks. This implementation intentionally estimates each stock separately so there is no accidental cross-stock pooling.

```{r eval=false}
independent_fit <- fit_independent_bayes(
  sim$panel,
  factor_names,
  chains = 2L,
  iter = 1500L,
  warmup = 750L,
  seed = 123L
)

independent <- extract_independent_betas(
  independent_fit,
  sim$panel,
  factor_names
)

independent_joined <- join_estimates_to_truth(
  independent,
  sim$truth$stock
)

independent_recovery <- summarise_beta_recovery(independent_joined)
independent_intervals <- summarise_interval_recovery(independent_joined)
```

## Hierarchical Bayesian fit

```{r eval=false}
hierarchical_fit <- fit_hierarchical_bayes(
  sim$panel,
  factor_names,
  chains = 4L,
  iter = 2000L,
  warmup = 1000L,
  seed = 123L,
  adapt_delta = 0.95,
  max_treedepth = 12L
)

hierarchical_diagnostics <- summarise_brms_diagnostics(
  hierarchical_fit,
  max_treedepth = 12L
)
hierarchical_diagnostics

hierarchical <- extract_hierarchical_betas(
  hierarchical_fit,
  sim$panel,
  factor_names
)

hierarchical_joined <- join_estimates_to_truth(
  hierarchical,
  sim$truth$stock
)

hierarchical_recovery <- summarise_beta_recovery(hierarchical_joined)
hierarchical_intervals <- summarise_interval_recovery(hierarchical_joined)
```

## Acceptance gate

Before comparing RMSE, require the hierarchical fit to satisfy the diagnostics gate.

```{r eval=false}
stopifnot(
  hierarchical_diagnostics$max_rhat < 1.01,
  hierarchical_diagnostics$divergences == 0,
  hierarchical_diagnostics$max_treedepth_hits == 0
)
```

ESS thresholds should be reviewed at parameter level rather than reduced to one arbitrary universal number.

## Common recovery table

```{r eval=false}
ols_recovery$estimator <- "OLS"
ridge_recovery$estimator <- "Ridge"
independent_recovery$estimator <- "Independent Bayes"
hierarchical_recovery$estimator <- "Hierarchical Bayes"

recovery <- rbind(
  ols_recovery,
  ridge_recovery,
  independent_recovery,
  hierarchical_recovery
)

recovery[order(recovery$factor, recovery$rmse), ]
```

## Shrinkage plot

The core visual should show the true beta, OLS estimate, ridge estimate, independent Bayesian posterior mean, and hierarchical posterior mean for the same stock-factor observations.

Do not cherry-pick only the most dramatic shrinkage examples. Plot all stocks for at least MKT, HML, and MOM, then provide the full six-factor table separately.

## Hierarchy-strength experiment

After the balanced case works, repeat the same recovery experiment under:

```{r}
hierarchy_scenarios()
```

Expected interpretation:

- `none`: sector identity should add little or no information;
- `weak`: hierarchical gains should be modest;
- `balanced`: partial pooling has a genuine opportunity to help;
- `strong`: sector pooling should be most useful, especially in short samples.

This expectation is a hypothesis, not a result.

## Wrong-hierarchy experiment

Then deliberately corrupt 25% and 50% of stock sector labels:

```{r}
corrupted_25 <- corrupt_sector_labels(sim$panel, 0.25, seed = 200)
corrupted_50 <- corrupt_sector_labels(sim$panel, 0.50, seed = 201)

sector_mapping_accuracy(corrupted_25)
sector_mapping_accuracy(corrupted_50)
```

When evaluating a corrupted hierarchy, match estimated betas to truth by stock and factor rather than by the deliberately wrong observed sector:

```{r eval=false}
wrong_hierarchy_fit <- fit_hierarchical_bayes(
  corrupted_25,
  factor_names,
  chains = 4L,
  iter = 2000L,
  seed = 202L
)

wrong_hierarchy_est <- extract_hierarchical_betas(
  wrong_hierarchy_fit,
  corrupted_25,
  factor_names
)

wrong_hierarchy_joined <- join_estimates_to_truth(
  wrong_hierarchy_est,
  sim$truth$stock,
  match_sector = FALSE
)
```

A credible project should show where hierarchical shrinkage becomes harmful when the economic grouping is wrong.

## Expansion order

Only after the 20-stock experiment is correct:

1. increase to 50 stocks;
2. increase to 100 stocks;
3. run the sample-size grid 26 / 52 / 104 / 156 / 260 / 520;
4. run hierarchy-strength scenarios;
5. run label corruption;
6. automate prior sensitivity;
7. then consider real data.
